---
title: "Test 1"
author: "Sirio Sobrino"
date: "29/9/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
```

## Data Analysis

1. Download the dataset xxx from [http://archive.ics.uci.edu/ml/datasets/Forest+Fires]
   + Read the csv and store in a data.table (raw_data), and check its structure and class 
   + Visualize the first 10 rows of raw_data and get a summary of each variable


```{r}
raw_data<-readRDS('AmesHousing_pro.RData')
raw_data<-raw_data[,.(SalePrice,GrLivArea,LotArea,GarageArea,FullBath,BedroomAbvGr,KitchenQual,Heating,PoolArea)]
str(raw_data)
head(raw_data, 10)
summary(raw_data)
```

2. Analize the correlation matrix of raw_data's numeric variables. Plot the pairwise correlations
   + Analize the presence of NAs in data. Which variable has any NAs?
   + Eliminate the NA value in variable with NAs and repeat the analysis, and repeat the correlation analysis
   

```{r}


var_numeric<-raw_data[,names(raw_data)[sapply(raw_data,is.numeric)],with=F]

apply(sapply(var_numeric,is.na),2,sum)


data.frame(cor(var_numeric))
pairs(var_numeric)

library(corrplot)
sum(is.na(var_numeric))
corrplot(cor(var_numeric),'number')

var_numeric<-var_numeric[!is.na(GarageArea)]
data.frame(cor(var_numeric))
pairs(var_numeric)
corrplot(cor(var_numeric),'number')

```


3. Choose the variable that best explains by itself the SalePrice and make a scatter plot
   + Calculate the linear correlation and store it in a new variable (cor_1)
   + Create a data.frame (df) with the cases of var_numeric with GrLivArea belonging to quantiles 0,1% and 99,9%
   + Repeat the correlation analysis (cor_2), and compare with cor_1; comment your results

```{r}

plot(var_numeric$GrLivArea, var_numeric$SalePrice); grid()
cor_1<-with(var_numeric,cor(GrLivArea,SalePrice))

lower_value<-quantile(var_numeric$GrLivArea,probs = 0.001)
upper_value<-quantile(var_numeric$GrLivArea,probs = 0.999)

df<-var_numeric[lower_value<GrLivArea&GrLivArea<upper_value]
plot(df$GrLivArea, df$SalePrice); grid()
cor_2<-with(df,cor(GrLivArea,SalePrice))

c(cor_1,cor_2)

```


4. Analize the variable PoolArea and create a new factor variable in raw_data (has_pool)
   + with values TRUE/FALSE if PoolArea is positive
   + analize numerically and grafically if having a pool increases the price
   + create a t.test to test if mean price is statistically different with pool and without it. Is it significant at 5% level of         significance? comment your results
   + repeat the t.test specifying tha the alternative is prices of houses with pool being greater than those of houses without pool. comment your results

```{r}
raw_data[,has_pool:=factor(ifelse(PoolArea>0,TRUE,FALSE))]

raw_data[,.(mean_price=mean(SalePrice)),by=has_pool]

library(ggplot2)

# boxplot comparative
qplot(x=raw_data$has_pool,y=raw_data$SalePrice,geom='boxplot')

# density comparative, in two separate plots
ggplot(raw_data,aes(x=SalePrice),colour=has_pool)+geom_density()+facet_wrap(~has_pool,nrow=2)


# density comparative, in one plot
data<-reshape2::melt(raw_data[,.(SalePrice,has_pool)])
ggplot(data,aes(x=value,fill=has_pool))+geom_density(alpha=0.5) 


help(t.test)
mean_pool_test<-t.test(raw_data[has_pool==TRUE]$SalePrice,raw_data[has_pool==FALSE]$SalePrice) # not significant at 95%
objects(mean_pool_test)
paste0("p_value = ",round(mean_pool_test$p.value,4))
paste0(c("lower 95% limit = ","upper 95% limit = "),round(mean_pool_test$conf.int,0))

# alternative:greater
mean_pool_test_2<-t.test(raw_data[has_pool==TRUE]$SalePrice,raw_data[has_pool==FALSE]$SalePrice, alternative = 'g') # it's significant at 95%
objects(mean_pool_test_2)
paste0("p_value = ",round(mean_pool_test_2$p.value,4))
paste0(c("lower 95% limit = ","upper 95% limit = "),round(mean_pool_test_2$conf.int,0))

```



