---
title: "Test 1"
author: "Sirio Sobrino"
date: "29/9/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Types

1. Create a numeric variable x with numbers 1 to 10

```{r}
x<-1:10
```

2. Create a new variable, y, equal to the square of x

```{r}
y<-x^2
```

3. Store both variables forming a matrix (m) with two columns.
   + Check the class of m
   + Print the first 10 rows of the m
   + Print the 4th element of y 

```{r}
m<-cbind(x,y)
class(m)
m[1:10,]
m[4,'y']
```

4. Pass from matrix to a dataframe object (df)
   + Add a new variable called z containing letters "A", and "B", repeated 5 times each
   + Check the structure of df and its variables classes

```{r}
df<-data.frame(m)
df$z<-c(rep('A',5),rep('B',5))
str(df)
```

5. Coerce the variable z to a factor variable
   + Check the structure of df and its variables classes

```{r}
df$z<-as.factor(df$z)
```



6. Subset the rows of df with numbers in x greater than 3 and z being "A", and store it in a new data.frame (df_1). Print it

```{r}
df_1<-df[df$x>3&df$z=="A",]
```


7. Generate a NA value in element 6th and 7th of y in df 
   + Create a new data.frame (df_2) from df filtering the NA's in y, and keeping only columns y and z. Print it
   + Get a summary of the variables in df_2

```{r}
df[6:7,'y']<-NA
df_2<-df[!is.na(df$y),c('y','z')]
df_2
```

8. Create a new data.frame (df_3) adding a new row to df_2, with values y=0, z="C". Print it
   + Create also a new data.frame df_4 with columns x and z from df

```{r}
df_3<-rbind(df_2, data.frame(y=0,z="C"))
df_3
str(df_3)

df_4<-df[,c('x','z')]
```


9. Create a new data.frame (df_5) from a subset of df_3 with y being 1, 100 or 0, 
    and another data.frame(df_6) from a subset of df_4 with x being 3 or 9 
   + Create a new data.frame (df_7) by merging df_6 with df_6 by variable z, but keeping all elements of df_5

```{r}
# df_3$z<-as.character(df_3$z)
# df_4$z<-as.character(df_4$z)

str(df_3)
str(df_4)

df_5<-df_3[df_3$y %in% c(1,100,0),]
df_5 

df_6<-df_4[df_4$x %in% c(3,9),]
df_6 

df_7<-merge(df_5,df_6, by='z', all.x=T)
df_7
str(df_7)
```


10. Create a list containing all the data types generated
    + Check the structure of the list
    + Save the list in your current wf as ar .RData file
    + Save the element 'm' of the list in a .csv file and in a .xlsx file
    
```{r}
output_list<-list(x,y,m,df,df_1,df_2,df_3,df_4,df_5,df_6,df_7)
str(output_list)

saveRDS(output_list,'output_list.RData')

write.csv(output_list[[3]],'output_list.csv')

require(xlsx)
# write.xlsx(output_list[[3]],'output_list.xlsx')

```
    