---
title: "08_Descriptve_Analysis"
author: "SirioSobrino"
date: "11/9/2017"
output: html_document
---

```{r setup, include=T}
knitr::opts_chunk$set(echo = T)
```

#### Summarizing a Variable: univariate and bivariate
#### Quantiles and Outlier Detection
#### Variance, Covariance and Linear Correlation

Lets's start donwloading the House Prices dataset from <http://rmarkdown.rstudio.com>. Then read it.

```{r raw_data_reading, echo=T}
.libPaths("/home/daadmin/R/x86_64-redhat-linux-gnu-library/3.2")
library(data.table)
library(knitr)
library(stargazer)
raw_data<-readRDS('AmesHousing_pro.RData')
```

## dataset summary

```{r raw_data_summary, echo=T}
str(raw_data)
if(length(unique(raw_data$PID))==nrow(raw_data)) print('PID is a unique index')
head(raw_data) 
knitr::kable(head(raw_data[,1:8,with=F],n=5))  # fancy table display
```

## single variable analysis

```{r single_variable, echo=T}

y<-as.numeric(raw_data$SalePrice) 
cat('y summary: ',summary(y))
quantile(y, seq(0,1,0.1))
y[1]
library(ggplot2)
library(gridExtra)

p0_y<-qplot(x=1:nrow(raw_data),y=y, geom='point')
p1_y<-qplot(y, geom='histogram')
p2_y<-qplot(y, geom='density')
p3_y<-qplot(y, x= 1, geom = "boxplot")
p_y<-list(p0_y,p1_y, p2_y, p3_y)
marrangeGrob(p_y, nrow=2, ncol=2)

# lets calculate the more important statistics of a variable: mean and standard deviation
y_mean<-mean(y)
y_sd<-sd(y)

# and calculate the Cooef. of Variation
print(paste0('Coeff. of Variation = ', 100*round(y_sd/y_mean, 2),'%'))

# outlier detection: lets mark every house with a price superior to 95% of all prices:
extreme_prices<- y>=quantile(y,0.95)
sum(extreme_prices)
sum(extreme_prices)/length(y)
plot(y, pch=19, xlab=''); grid()
points(y*as.numeric(extreme_prices),col='red')
points(rep(0,length(y)),col='white')

# As we can derive from the univariate analysis, sale prices are highly 
# right-skeed
# the smallest the variance, the more representative becomes the mean from the population, but 
# the distribution should ideally be centered

# lets try taking logarithms on sale price

ln_y<-log(y)

p0_ln<-qplot(x=1:nrow(raw_data),y=ln_y, geom='point')
p1_ln<-qplot(ln_y, geom='histogram')
p2_ln<-qplot(ln_y, geom='density')
p3_ln<-qplot(ln_y, x= 1, geom = "boxplot")
p_ln<-list(p0_ln,p1_ln,p2_ln,p3_ln)
marrangeGrob(p_ln, nrow=2, ncol=2)

```


## by-variate analysis

lets take the living area surface (square feet) and sale price (dolars) and analize their relationship

```{r bi_variable, echo=T}

df<-raw_data[,.(PID,x=GrLivArea, y=SalePrice)]
head(df)

p1<-qplot(df$x,geom='density')
p2<-qplot(df$y,geom='density')
p4<-ggplot(df,aes(x=x,y=y))+geom_point()
p<-list(p1,p2)
marrangeGrob(p, nrow=2, ncol=1)
print(p4+ggtitle('Living Area Surface Vs Price'))

with(df,cov(x,y))  

with(df,cor(x,y))
attach(df)
cor(x,y)==cov(x,y)/(sd(x)*sd(y))

lm(y~x)$coef['x']
detach(df)

```

We can also display interactive charts based on ggplot, via the ggplotly package
```{r interactive_plot,echo=T, message=F}

library(plotly)

ggplotly(ggplot(df,aes(x=x,y=y))+geom_point(),
         width = 600, height = 400)
ggplotly(ggplot(df,aes(x=x,y=y))+geom_point()+stat_smooth(method='auto'),
         width = 600, height = 400)

```



Correlation Matrix: a matrix indicating all pairwise linear correlation of a set of variables

```{r correlation_matrix, echo=T}
library(corrplot)

data_sample<-raw_data[,.(SalePrice,
                         GrLivArea,
                         GarageArea,
                         TotalBsmtSF,
                         BsmtUnfSF,
                         BsmtFinSF1,
                         LotArea)]
head(data_sample)
corrplot(cor(data_sample[complete.cases(data_sample)]), method = 'number', tl.col = 'black')

library(GGally)

p<-GGally::ggpairs(data_sample[complete.cases(data_sample)],axisLabels = 'none',size=1,lwd=0.5)
ggplotly(p, width = 800, height = 500)

```